\Large {\bf Protocol-Guided Analysis of Post-silicon Traces Under Limited Observability}

### 2015-09-12 paper.tex
revised section II
Figure 1 and 2 are merged such that a LPN is created for the BPMN in
figure 1.  The last paragraph of section II is also revised accordingly.

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
More specifically, given a collection of system-level
communication protocols and a trace of (partially observed)
hardware signals, our approach infers, with a certain
measure of confidence, the protocol instances (and their
interleavings) being exercised by the trace.  Our approach
is based on a formalization of system-level transactions via
labeled Petri-Nets~\cite{petrinets}, which are capable of
describing sequencing, concurrency, and choices of events.
Given this formalization, we develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values, together with an estimate of
confidence on the inference.  We demonstrate our approach on
two SoC models: a SystemC prototype and a realistic
implementation constructed within the GEM5
environment~\cite{Binkert2011}.

@@

@@

An SoC design is typically composed of a large number of
pre-designed hardware or software blocks (often referred to
as ``intellectual properties'' or ``IPs'') that coordinate
through complex protocols to implement system-level
behavior.  Fig.~\ref{fig:flow-spec-ex} shows one such
protocol, involving th
\begin{figure}[tb]
\begin{center}
\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
\label{flow-spec-ex}
\end{center}
\end{figure}

@@

Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.  In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.  Labeled Petri nets are capable of describing sequencing, concurrency, and choices.   This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.

@@

A simple communication protocol specification between two components represented by a LPN is shown in~Figure~\ref{fig-flow-ex-1}.  In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.  Each transition is labeled with its name and the asSoCiated event.  The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.  In this LPN, the initial state $\mathit{init} = \{p_1\}$.  In this protocol, component~1 sends two consecutive messages $\mathit{msg}_1$ to component~2, which responds with either $\mathit{msg}_2$ or $\mathit{msg}_3$ non-deterministically.  If $\mathit{msg}_2$ is responded, component~1 repeats the protocol by sending two consecutive $\mathit{msg}_1$ again.  Otherwise, component~1 stops the protocol upon receiving $\mathit{msg}_3$.  This LPN represents two possible communication sequences as shown below.
\[
\begin{array}{l}
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\ldots \\
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\ldots, \mathit{msg}_3\;
\end{array}
\]

\begin{figure}[tb]
\begin{center}
\resizebox{2.5in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:msg_1$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:msg_2$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (3,0)$) {$t_3:msg_3$};
	\node[rectbox]	(t4)	at	($(p3) + (-3,0)$) {$t_4:msg_4$};
	\node[circ,fill=black] 	(term) 		at ($(t4) + (0,-1.1)$) {$~~~$};

	% Edges
	\draw[->,weight3]	(p1) -- (t1);
	\draw[->,weight3]	(t1) -- (p2);
	\draw[->,weight3]	(p2) -- (t2);
	\draw[->,weight3]	(t2) -- (p3);
	\draw[->,weight3]	(p3) -- (t3);
	\draw[->,weight3]	(p3) -- (t4);
	\draw[->,weight3]	(t4) -- (term);	
	\draw[->,weight3]	(t3) -- ($(p1)+(3,0)$) -- (p1);

\end{tikzpicture}
}
\caption{An example of a LPN representing a system flow.}
\label{fig-flow-ex-1}
\end{center}
\end{figure}

@@

% @Note  \red{\bf Another assumption:  the observed signal trace captures the values on the observed signals in every cycle from the beginning.  Too strong an assumption??}

>>>>

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
More specifically, given a collection of system-level
communication protocols and a trace of (partially observed)
hardware signals, our approach infers, with a certain
measure of confidence, the protocol instances (and their
interleavings) being exercised by the trace.  Our approach
is based on a formalization of system-level transactions via
labeled Petri-Nets, which are capable of
describing sequencing, concurrency, and choices of events.
Given this formalization, we develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values, together with an estimate of
confidence on the inference.  We demonstrate our approach on
two SoC models: a SystemC prototype and a more realistic
implementation constructed within the GEM5
environment~\cite{Binkert2011}.

@@

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};

	% Edges
	\draw[->,weight3]	(p1) -- (t1);
	\draw[->,weight3]	(t1) -- (p2);
	\draw[->,weight3]	(p2) -- (t2);
	\draw[->,weight3]	(t2) -- (p3);
	\draw[->,weight3]	(p3) -- (t3);
	\draw[->,weight3]	(t3) -- (p4);
	\draw[->,weight3]	(p4) -- ($(p4)+(-3,0)$) -- (t4);
	\draw[->,weight3]	(p4) -- ($(p4)+(3,0)$) -- (t5);
	\draw[->,weight3]	(t4) -- (term1);	
	\draw[->,weight3]	(t5) -- (term2);

\end{tikzpicture}
}
\end{minipage}
\\
(a) & (b)
\end{tabular}
\caption{(a) A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}, and (b) the LPN capturing the communication protocol in (a).}
\label{flow-spec-ex}
\end{center}
\end{figure*}

@@

An SoC design is typically composed of a large number of
pre-designed hardware or software blocks (often referred to
as ``intellectual properties'' or ``IPs'') that coordinate
through complex protocols to implement system-level
behavior.  Fig.~\ref{fig:flow-spec-ex} shows one such
protocol, involving th

@@

Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.  In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.  Labeled Petri nets, which are a subclass of traditional Petri-nets~\cite{Murata89} where Petri-net transitions are labeled with user defined information,  are capable of describing sequencing, concurrency, and choices.   This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.

@@

The communication protocol in the BMPN shown in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN shown in~Figure~\ref{flow-spec-ex}(b).  In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.  Each transition is labeled with its name and the associated event.  Each event has a form of $\langle {\tt src, dest, cmd} \rangle$ where ${\tt cmd}$ is a command sent from a source component ${\tt src}$ to a destination component ${\tt dest}$. The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.  The initial marking is $\mathit{init} = \{p_1\}$.  In this LPN model, only the communication portion of the BPMN specification is represented while the computation portion is ignored.

@@

\red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}
